---
title: "Data Visualization Part 1"
output: github_document
---

Initial Libraries
```{r}
library(tidyverse)
library(ggridges)

#Global Markdown Knitting Settings
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

Importing Data from NOAA Online Server

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

# Making a Plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + geom_point()

```

Identifying it through Piping

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_nyc_weather = 
  weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_nyc_weather

```

# Editing Plots and Aesthetic Mapping

```{r}
ggplot(weather_df, aes(x= tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth(se = FALSE)

#color maps onto points, smooth creates a curve to plot into the line
```

Plot with Facets 

```{r}
ggplot(weather_df, aes(x = tmin, y= tmax, color = name)) +
  geom_point(alpha = .3) + 
  geom_smooth() +
  facet_grid(. ~ name)

```

New Plot 

```{r}
ggplot(weather_df, aes(x= date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .3) +
  geom_smooth() +
  facet_grid(. ~ name)
```

Try assigning a specific color

```{r}
weather_df |> 
  filter(name != "CentralPark_NY") |> 
  ggplot(aes(x= date, y = tmax, color = name)) +
  geom_point(alpha = .7, size = .5)


```

If you want to assign color, best to put it in the geom_point for a particular trend curve
Easiest to filter trend of interest ahead of time

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x= date, y = tmax)) +
  geom_point(alpha = .7, size = .5, color = "blue")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()
```

```{r}
weather_df |> 
  filter(name == "Molokai_HI") |> 
  ggplot(aes(x = date, y = tmax)) +
  geom_line(alpha = 0.5, color = "green") +
  geom_point(size = 0.5)
```


## Univariate Plotting

Histogram

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")

# with these specifications on position not very readable
```

Let's use a density plot

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_density(alpha = .3, adjust = 0.7)
```

Using Boxplots

```{r}
ggplot(weather_df, aes(y = tmax, x = name)) + 
  geom_boxplot()
```

Violin Plot
```{r}
ggplot(weather_df, aes(y = tmax, x = name, fill = name)) + 
  geom_violin()
```

Ridge Plot
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) +
  geom_density_ridges()

```

# Saving and Embedding Plots
Setting a folder directory ahead of time in repositor

```{r}
ggp_weather = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_weather

ggsave("Results/ggp_weather.pdf", ggp_weather)
```

Resizing a figure within specific code chunk for embedding in Markdown File
```{r, fig.width = 12, fig.height = 12}
ggp_weather
```

```{r, fig.width = 6, fig.asp = .6}
ggp_weather
```

